---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r Load data, echo=TRUE, warning=FALSE}
unzip("activity.zip")
activity <- read.csv("activity.csv", stringsAsFactors = FALSE)
activity$date <- as.Date(activity$date)
head(activity)
```

## What is mean total number of steps taken per day?
```{r mean steps per day, echo=TRUE, warning=FALSE, message=FALSE}
library(dplyr)
stepsperday <- activity %>% group_by(date) %>% summarize(steps=sum(steps,na.rm = TRUE))
head(stepsperday)

hist(stepsperday$steps, main = "Histogram of total steps per day", xlab = 'Steps/day')
```

The mean total steps is `r format(mean(stepsperday$steps), big.mark = ',')` and median total steps is `r format(median(stepsperday$steps), big.mark = ',')`


## What is the average daily activity pattern?
```{r daily activity pattern, warning=FALSE, message=FALSE}
library(ggplot2)

stepsperinterval <- activity %>% group_by(interval) %>% summarise(steps = mean(steps, na.rm = TRUE))

mostActiveInterval = stepsperinterval[which.max(stepsperinterval$steps), ]$interval

ggplot(stepsperinterval, aes(x= interval, y=steps)) + geom_line()

```

On average `r paste(mostActiveInterval%/%100, mostActiveInterval%%100, '-', mostActiveInterval%/%100, (mostActiveInterval%%100) +5 )`  has most number of steps during the study period

## Imputing missing values
We will impute missing values with average number of steps during that interval rounded off to nearest integer.
```{r imupting values}
avgStepsInInterval <- function (interval, stepsperinterval) {
  round(stepsperinterval[stepsperinterval$interval == interval,]$steps)
}

imputedactivity <- activity %>% mutate(steps = ifelse(is.na(steps), avgStepsInInterval(interval, stepsperinterval), steps))

imputedstepsperday <- imputedactivity %>% group_by(date) %>% summarize(steps=sum(steps,na.rm = TRUE))
hist(imputedstepsperday$steps, main = "Histogram of total steps per day after imputing", xlab = 'Steps/day')

```

The mean total steps now is `r format(mean(imputedstepsperday$steps), big.mark = ',')` and median total steps is `r format(median(imputedstepsperday$steps), big.mark = ',')`. The histograms look similar and the means, medians and total steps have slightly increased. 

## Are there differences in activity patterns between weekdays and weekends?
```{r weekday weekends}
imputedactivity <- imputedactivity %>% mutate(day = factor(ifelse(weekdays(date) %in% c('Sunday','Saturday'), 'Weekend', 'Weekday')))

imputedstepsperinterval <- imputedactivity %>% group_by(interval, day) %>% summarise(steps = mean(steps, na.rm = TRUE))


ggplot(imputedstepsperinterval, aes(x= interval, y=steps)) + geom_line() +
  facet_grid(day~.)

```

We can see during weekdays our subject starts early and rushes to work in the morning and is only modaretely active in the evening threafter whereas in the weekends, sleeps slightly more and is fairly active throughout the day.
