---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r Load data, echo=TRUE}
#unzip("activity.zip")
activity <- read.csv("activity.csv", stringsAsFactors = FALSE)
activity$date <- as.Date(activity$date)
head(activity)
```

## What is mean total number of steps taken per day?
```{r mean steps per day, echo=TRUE}
library(dplyr)
stepsperday <- activity %>% group_by(date) %>% summarize(totalsteps=sum(steps,na.rm = TRUE))
head(stepsperday)
hist(stepsperday$totalsteps, main = "Histogram of total steps per day", xlab = 'Steps/day')
```

The mean total steps is `r format(mean(stepsperday$totalsteps), big.mark = ',')` and median total steps is `r format(median(stepsperday$totalsteps), big.mark = ',')`



## What is the average daily activity pattern?
```{r daily activity pattern}
library(ggplot2)
stepsperinterval <- activity %>% group_by(interval) %>% summarise(steps = mean(steps, na.rm = TRUE))
ggplot(stepsperinterval, aes(x= interval, y=steps)) + geom_line() + theme_classic()
mostActiveInterval = stepsperinterval[which.max(stepsperinterval$steps), ]$interval
```

On average `r paste(mostActiveInterval%/%100, mostActiveInterval%%100, '-', mostActiveInterval%/%100, (mostActiveInterval%%100) +5 )`  has most number of steps during the study period

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
